### 2015-06-03 SobolIndices.tex
Some examples for Sobol indices

\documentclass[]{amsart}

\usepackage{array,colortbl}
\usepackage{amsmath,amsfonts,amssymb,bm,listings} % no amsthm, Springer defines Theorem, Lemma, etc themselves
\usepackage{xspace}
\usepackage[autolinebreaks]{mcode}
\input FJHDef
\newcommand{\oerr}{e}

\begin{document}
\title{Adaptive Computation of Sobol' Indices}
\maketitle

\noindent To:  Cl\'ementine Prieur (U. Josef Fourier), Llu\'is Antoni Jim\'enez Rugama (IIT)

\bigskip

Today Cl\'ementine discussed the possibility of using \texttt{cubSobol\_g} to calculate Sobol sensitivity indices for functions $f : [0,1]^d \to \reals$.  Here is some sample code where we showed proof of concept.

\lstinputlisting{ProgramsImages/SobolIndexExample.m}

There are a few next steps:

\begin{enumerate}
\item Sobol' indices can be written in terms of $2d-1$-dimensional integrals.   Cl\'ementine will work out what those integrals are.

\item The problem that we want to solve is to find $s_1, \ldots, s_p$ such that 
\[
\max_{j=1, \ldots, p} \abs{S_j(I_1(f), \ldots, I_q(f)) - s_j} \le \varepsilon,
\]
 where the $I_\ell$ are integrals of $f$ or maybe powers of $f$ or something that Cl\'ementine will work out.  Tony and I need to think about how to modify \texttt{cubSobol\_g}.

My suggestion is that for each $m = m_{\min}, m_{\min}+1, \ldots$ we compute our data-based error bounds for the Sobol' cubatures $\hI_{\ell,m}(f)$ using $n=2^m$ points, i.e., 
\[
\abs{I_\ell(f) - \hI_{\ell,m}(f)} \le \oerr_{\ell,m}(f).
\]
We also identify $\hS$ such that 
\[
\hS_j(b_1, \ldots, b_q, c_1, \ldots, c_q) \le S(a_1, \ldots, a_q) \le \hS_j(c_1, \ldots, c_q, b_1, \ldots, b_q) 
\]
for all $b_\ell \le a_\ell \le c_\ell$.  When we get 
\begin{multline*}
\hS_j(\hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f), \ldots \\
\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f)) \\
 - \hS_j(\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f), \ldots \\
 \hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f))   \le 2\varepsilon, \quad j = 1, \ldots, p
\end{multline*}
we may stop and choose 
\begin{multline*}
s_j = \frac{1}{2}\Bigl[\hS_j(\hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f), \ldots \\
\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f)) \\
+ \hS_j(\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f), \ldots \\
 \hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f)) \Bigr]
 \end{multline*}

\item Cl\'ementine will talk to her student to involve him in this project.

\end{enumerate}

\bigskip

Any questions, clarifications, or comments?

\end{document}