### 2015-06-09 SobolIndices.tex
Update SobolIndices.tex

\documentclass[]{amsart}

\usepackage{array,colortbl}
\usepackage{amsmath,amsfonts,amssymb,bm,listings} % no amsthm, Springer defines Theorem, Lemma, etc themselves
\usepackage{xspace}
\usepackage[autolinebreaks]{mcode}
\input FJHDef
\newcommand{\oerr}{e}

\begin{document}
\title{Adaptive Computation of Sobol' Indices}
\maketitle

\noindent To:  Cl\'ementine Prieur (U. Josef Fourier), Llu\'is Antoni Jim\'enez Rugama (IIT)

\bigskip

Today Cl\'ementine discussed the possibility of using \texttt{cubSobol\_g} to calculate Sobol sensitivity indices for functions $f : [0,1]^d \to \reals$.  Here is some sample code where we showed proof of concept.

\lstinputlisting{ProgramsImages/SobolIndexExample.m}

There are a few next steps:

\begin{enumerate}
\item Sobol' indices can be written in terms of $2d-1$-dimensional integrals.   Cl\'ementine will work out what those integrals are.

\item The problem that we want to solve is to find $s_1, \ldots, s_p$ such that 
\[
\max_{j=1, \ldots, p} \abs{S_j(I_1(f), \ldots, I_q(f)) - s_j} \le \varepsilon,
\]
 where the $I_\ell$ are integrals of $f$ or maybe powers of $f$ or something that Cl\'ementine will work out.  Tony and I need to think about how to modify \texttt{cubSobol\_g}.

My suggestion is that for each $m = m_{\min}, m_{\min}+1, \ldots$ we compute our data-based error bounds for the Sobol' cubatures $\hI_{\ell,m}(f)$ using $n=2^m$ points, i.e., 
\[
\abs{I_\ell(f) - \hI_{\ell,m}(f)} \le \oerr_{\ell,m}(f).
\]
We also identify $\hS$ such that 
\[
\hS_j(b_1, \ldots, b_q, c_1, \ldots, c_q) \le S(a_1, \ldots, a_q) \le \hS_j(c_1, \ldots, c_q, b_1, \ldots, b_q) 
\]
for all $b_\ell \le a_\ell \le c_\ell$.  When we get 
\begin{multline*}
\hS_j(\hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f), \ldots \\
\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f)) \\
 - \hS_j(\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f), \ldots \\
 \hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f))   \le 2\varepsilon, \quad j = 1, \ldots, p
\end{multline*}
we may stop and choose 
\begin{multline*}
s_j = \frac{1}{2}\Bigl[\hS_j(\hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f), \ldots \\
\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f)) \\
+ \hS_j(\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f), \ldots \\
 \hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f)) \Bigr]
 \end{multline*}

\item Cl\'ementine will talk to her student to involve him in this project.

\end{enumerate}

$$\mu=\mathbb{E}\left[f(\underline{X})\right]=\int_{[0,1]^d} f(\underline{x}) d \underline{x} \, .$$
$$\Var \left(f(\underline{X})\right)=\int_{[0,1]^d}\left(f(\underline{x})-\mu\right)^2d \underline{x} =\int_{[0,1]^d}\left(f(\underline{x})\right)^2d \underline{x} - \mu^2 \, .$$
$$\mu_j(x)=\mathbb{E}\left[f(\underline{X})|X_j=x\right]=\int_{[0,1]^{d-1}} f(\underline{t}_{-j},x)d\underline{t}_{-j} \, .$$
$$\begin{array}{rcl}
\mathbb{E}\left\{ \Var \left(f(\underline{X})|X_j\right)\right\} & = &\int_{0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_0^1 \mu_j^2(x) dx  \\
&=& \int_{[0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_{[0,1]^{2d-1}}f(\underline{x})f(\underline{x}'_{-j},x_j)d \underline{x} d \underline{x}'_{-j} \, . 
\end{array}$$

$$\mu_{-j}(\underline{x}_{-j})=\mathbb{E}\left[f(\underline{X})|X_{-j}=\underline{x}_{-j}\right]=\int_0^1 f({t}_{j},\underline{x}_{-j})d {t}_{j} \, .$$
$$\begin{array}{rcl}
\mathbb{E}\left\{ \Var \left(f(\underline{X})|\underline{X}_{-j}\right)\right\} & = &\int_{[0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_{[0,1]^{d-1}} \mu_{-j}^2(\underline{x}_{-j}) d \underline{x}_{-j}  \\
&=& \int_{[0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_{[0,1]^{d+1}}f(\underline{x})f(x'_{j},\underline{x}_{-j})d \underline{x} d {x}'_{j} \, . 
\end{array}$$

\bigskip

$$S_{j}=1-\frac{\mathbb{E}\left\{ \Var \left(f(\underline{X})|X_j\right)\right\}}{\Var\left(f(\underline{X})\right)}= \frac{N_j}{D}\, .$$
$$S_{j}^{\textup{tot}}=\frac{\mathbb{E}\left\{ \Var \left(f(\underline{X})|\underline{X}_{-j}\right)\right\} }{\Var\left(f(\underline{X})\right)}=\frac{N_j^{\textup{tot}}}{D}\, .$$

As $N_j$, $N_j^{\textup{tot}}$ and $D$ can be written as sums of integrals,
one can guarantee with the existing code the existence of error terms $\tilde{\oerr}_{j,m}(f)$, $\tilde{\oerr}_{j,m}^{\textup{tot}}(f)$ and $\tilde{\oerr}_m(f)$ such that

$$\hat{N}_{j,m}-\tilde{\oerr}_{j,m}(f) \leq N_j \leq \hat{N}_{j,m}+\tilde{\oerr}_{j,m}(f)  \, ,$$
$$\hat{N}_{j,m}^{\textup{tot}}-\tilde{\oerr}_{j,m}^{\textup{tot}}(f) \leq N_j^{\textup{tot}} \leq \hat{N}_{j,m}^{\textup{tot}}+\tilde{\oerr}_{j,m}^{\textup{tot}}(f) \, ,$$
$$\hat{D}_{m}-\tilde{\oerr}_{m}(f) \leq D \leq \hat{D}_{m}+\tilde{\oerr}_{m}(f)  \, .$$

Thus

$$\frac{\hat{N}_{j,m}-\tilde{\oerr}_{j,m}(f)}{\hat{D}_{m}+\tilde{\oerr}_{m}(f) } \leq \frac{N_j}{D} \leq \frac{\hat{N}_{j,m}+\tilde{\oerr}_{j,m}(f)}{\hat{D}_{m}-\tilde{\oerr}_{m}(f) }\, ,$$
$$ \frac{\hat{N}_{j,m}^{\textup{tot}}-\tilde{\oerr}_{j,m}^{\textup{tot}}(f) }{\hat{D}_{m}+\tilde{\oerr}_{m}(f)} \leq \frac{N_{j}^{\textup{tot}}}{D} \leq \frac{ \hat{N}_{j,m}^{\textup{tot}}+\tilde{\oerr}_{j,m}^{\textup{tot}}(f)}{\hat{D}_{m}-\tilde{\oerr}_{m}(f)} \, .$$

\bigskip

Any questions, clarifications, or comments?

\end{document}