### 2015-07-16 SobolIndices.tex
Adding some more notes and starting code for the computations.

\documentclass[]{amsart}

\usepackage{array,colortbl}
\usepackage{amsmath,amsfonts,amssymb,bm,listings} % no amsthm, Springer defines Theorem, Lemma, etc themselves
\usepackage{xspace}
\usepackage[autolinebreaks]{mcode}
\input FJHDef
\newcommand{\oerr}{e}

\begin{document}
\title{Adaptive Computation of Sobol' Indices}
\maketitle

\noindent To:  Cl\'ementine Prieur (U. Josef Fourier), Llu\'is Antoni Jim\'enez Rugama (IIT)

\bigskip

Today Cl\'ementine discussed the possibility of using \texttt{cubSobol\_g} to calculate Sobol sensitivity indices for functions $f : [0,1]^d \to \reals$.  Here is some sample code where we showed proof of concept.

\lstinputlisting{ProgramsImages/SobolIndexExample.m}

There are a few next steps:

\begin{enumerate}
\item Sobol' indices can be written in terms of $2d-1$-dimensional integrals.   Cl\'ementine will work out what those integrals are.

\item The problem that we want to solve is to find $s_1, \ldots, s_p$ such that 
\[
\max_{j=1, \ldots, p} \abs{S_j(I_1(f), \ldots, I_q(f)) - s_j} \le \varepsilon,
\]
 where the $I_\ell$ are integrals of $f$ or maybe powers of $f$ or something that Cl\'ementine will work out.  Tony and I need to think about how to modify \texttt{cubSobol\_g}.

My suggestion is that for each $m = m_{\min}, m_{\min}+1, \ldots$ we compute our data-based error bounds for the Sobol' cubatures $\hI_{\ell,m}(f)$ using $n=2^m$ points, i.e., 
\[
\abs{I_\ell(f) - \hI_{\ell,m}(f)} \le \oerr_{\ell,m}(f).
\]
We also identify $\hS$ such that 
\[
\hS_j(b_1, \ldots, b_q, c_1, \ldots, c_q) \le S(a_1, \ldots, a_q) \le \hS_j(c_1, \ldots, c_q, b_1, \ldots, b_q) 
\]
for all $b_\ell \le a_\ell \le c_\ell$.  When we get 
\begin{multline}\label{stopping_crit}
\hS_j(\hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f), \ldots \\
\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f)) \\
 - \hS_j(\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f), \ldots \\
 \hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f))   \le 2\varepsilon, \quad j = 1, \ldots, p
\end{multline}
we may stop and choose 
\begin{multline*}
s_j = \frac{1}{2}\Bigl[\hS_j(\hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f), \ldots \\
\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f)) \\
+ \hS_j(\hI_{1,m}(f) - \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) - \oerr_{q,m}(f), \ldots \\
 \hI_{1,m}(f) + \oerr_{1,m}(f), \ldots,\hI_{q,m}(f) + \oerr_{q,m}(f)) \Bigr]
 \end{multline*}

\item Cl\'ementine will talk to her student to involve him in this project.

\end{enumerate}

$$\mu=\mathbb{E}\left[f(\underline{X})\right]=\int_{[0,1]^d} f(\underline{x}) d \underline{x} \, .$$
$$\Var \left(f(\underline{X})\right)=\int_{[0,1]^d}\left(f(\underline{x})-\mu\right)^2d \underline{x} =\int_{[0,1]^d}\left(f(\underline{x})\right)^2d \underline{x} - \mu^2 \, .$$
$$\mu_j(x)=\mathbb{E}\left[f(\underline{X})|X_j=x\right]=\int_{[0,1]^{d-1}} f(\underline{t}_{-j},x)d\underline{t}_{-j} \, .$$
$$\begin{array}{rcl}
\mathbb{E}\left\{ \Var \left(f(\underline{X})|X_j\right)\right\} & = &\int_{0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_0^1 \mu_j^2(x) dx  \\
&=& \int_{[0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_{[0,1]^{2d-1}}f(\underline{x})f(\underline{x}'_{-j},x_j)d \underline{x} d \underline{x}'_{-j} \, . 
\end{array}$$

$$\mu_{-j}(\underline{x}_{-j})=\mathbb{E}\left[f(\underline{X})|X_{-j}=\underline{x}_{-j}\right]=\int_0^1 f({t}_{j},\underline{x}_{-j})d {t}_{j} \, .$$
$$\begin{array}{rcl}
\mathbb{E}\left\{ \Var \left(f(\underline{X})|\underline{X}_{-j}\right)\right\} & = &\int_{[0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_{[0,1]^{d-1}} \mu_{-j}^2(\underline{x}_{-j}) d \underline{x}_{-j}  \\
&=& \int_{[0,1]^d}\left[f(\underline{x}\right]^2 d \underline{x} - \int_{[0,1]^{d+1}}f(\underline{x})f(x'_{j},\underline{x}_{-j})d \underline{x} d {x}'_{j} \, . 
\end{array}$$

\bigskip

$$S_{j}=1-\frac{\mathbb{E}\left\{ \Var \left(f(\underline{X})|X_j\right)\right\}}{\Var\left(f(\underline{X})\right)}= \frac{N_j}{D}\, .$$
$$S_{j}^{\textup{tot}}=\frac{\mathbb{E}\left\{ \Var \left(f(\underline{X})|\underline{X}_{-j}\right)\right\} }{\Var\left(f(\underline{X})\right)}=\frac{N_j^{\textup{tot}}}{D}\, .$$

As $N_j$, $N_j^{\textup{tot}}$ and $D$ can be written as sums of integrals,
one can guarantee with the existing code the existence of error terms $\tilde{\oerr}_{j,m}(f)$, $\tilde{\oerr}_{j,m}^{\textup{tot}}(f)$ and $\tilde{\oerr}_m(f)$ such that

$$\hat{N}_{j,m}-\tilde{\oerr}_{j,m}(f) \leq N_j \leq \hat{N}_{j,m}+\tilde{\oerr}_{j,m}(f)  \, ,$$
$$\hat{N}_{j,m}^{\textup{tot}}-\tilde{\oerr}_{j,m}^{\textup{tot}}(f) \leq N_j^{\textup{tot}} \leq \hat{N}_{j,m}^{\textup{tot}}+\tilde{\oerr}_{j,m}^{\textup{tot}}(f) \, ,$$
$$\hat{D}_{m}-\tilde{\oerr}_{m}(f) \leq D \leq \hat{D}_{m}+\tilde{\oerr}_{m}(f)  \, .$$

Thus

\begin{equation}\label{err_bound_j}
\frac{\hat{N}_{j,m}-\tilde{\oerr}_{j,m}(f)}{\hat{D}_{m}+\tilde{\oerr}_{m}(f) } \leq \frac{N_j}{D} \leq \frac{\hat{N}_{j,m}+\tilde{\oerr}_{j,m}(f)}{\hat{D}_{m}-\tilde{\oerr}_{m}(f) }\, ,
\end{equation}
\begin{equation}\label{err_bound_jtot}
\frac{\hat{N}_{j,m}^{\textup{tot}}-\tilde{\oerr}_{j,m}^{\textup{tot}}(f) }{\hat{D}_{m}+\tilde{\oerr}_{m}(f)} \leq \frac{N_{j}^{\textup{tot}}}{D} \leq \frac{ \hat{N}_{j,m}^{\textup{tot}}+\tilde{\oerr}_{j,m}^{\textup{tot}}(f)}{\hat{D}_{m}-\tilde{\oerr}_{m}(f)} \, .
\end{equation}

\bigskip

Any questions, clarifications, or comments?

Domains for integrals $N_j$ and $N_j^{\textup{tot}}$ can be rearranged to consider them as a single integral:
\begin{align*}
\mathbb{E}\left[ \Var \left(f(\underline{X})|X_j\right)\right] & = \int_{[0,1)^d}f(\underline{x})^2 d \underline{x} - \int_{[0,1)^{2d-1}}f(\underline{x})f(\underline{x}'_{-j},x_j)d \underline{x} d \underline{x}'_{-j} \\
& = \int_{[0,1)^{2d-1}}\left(f(\underline{x})^2-f(\underline{x})\right)f(\underline{x}'_{-j},x_j)d \underline{x} d \underline{x}'_{-j} \, . 
\end{align*}

\begin{align*}
\mathbb{E}\left[ \Var \left(f(\underline{X})|\underline{X}_{-j}\right)\right] & = \int_{[0,1)^d}f(\underline{x})^2 d \underline{x} - \int_{[0,1)^{d+1}}f(\underline{x})f(x'_{j},\underline{x}_{-j})d \underline{x} d {x}'_{j} \\
& = \int_{[0,1)^{d+1}}\left(f(\underline{x})^2-f(\underline{x})\right)f(x'_{j},\underline{x}_{-j})d \underline{x} d {x}'_{j} \, . 
\end{align*}

However, because $D$ is a nonlinear combination of integrals over different domains, it needs to be split into 2 parts:

\begin{align*}
\Var \left(f(\underline{X})\right) & = \int_{[0,1)^d}f(\underline{x})^2 d \underline{x} - \left(\int_{[0,1)^d}f(\underline{x}) d \underline{x}\right)^2 \\
& = D_1 - D_2^2 \, .
\end{align*}

Namely $a_1^{(j)} :=\mathbb{E}\left[ \Var \left(f(\underline{X})|X_j\right)\right]$, $a_2^{(j)} := \mathbb{E}\left[ \Var \left(f(\underline{X})|\underline{X}_{-j}\right)\right]$, $a_3^{(j)} := D_1$ and $a_4^{(j)} := D_2$ (we will spare the ${(j)}$ notation for simplicity), we have:

\begin{equation}
S(a_1,\dots,a_4)=1-\frac{a_1}{a_3-a_4^2} \; , \qquad S^\textup{tot}(a_1,\dots,a_4)=\frac{a_2}{a_3-a_4^2} \, .
\end{equation}
that directly leads to,
\begin{equation}\label{functionals_stop_criteria}
\hS(b_1,\dots,b_4,c_1,\dots,c_4)=1-\frac{c_1}{b_3-c_4^2} \; , \qquad \hS^\textup{tot}(b_1,\dots,b_4,c_1,\dots,c_4)=\frac{b_2}{c_3-b_4^2} \, .
\end{equation}
Remark that if we have some a priori knowledge of $a_4$, $c_4^2$ can be replaced by $c_4$ and $b_4^2$ by $b_4$ depending on whether $a_4 \leq 1$ or $a_4 \geq 1$ respectively. This would provide tighter bounds.

According to \eqref{functionals_stop_criteria},
\begin{equation}
\hS^\textup{tot}(b_1,b_2,b_3,b_4,c_1,c_2,c_3,c_4) = 1 - \hS(b_1,c_1,c_3,c_4,b_2,c_2,b_3,b_4) \, .
\end{equation}
Hence, we do not really need a second estimator $\hS^\textup{tot}$. For further clarity, we can also set to $0$ the variables not affecting our estimators, $\hS(b_1,0,b_3,b_4,c_1,0,c_3,c_4)$ and $1-\hS(0,0,c_3,c_4,b_2,c_2,b_3,b_4)$.

From here, we can use the stopping criteria proposed in \eqref{stopping_crit}.

\end{document}