### 2016-09-28 article/Images/boxplots.tex
changes L, section 5, abstract, conclusion, MSC codes

% Created by tikzDevice version 0.10.1 on 2016-09-27 19:09:28
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,361.35);
\begin{scope}
\path[clip] ( 52.08, 52.08) rectangle (416.82,344.55);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.2pt,line join=round] ( 81.18,261.99) -- (101.96,261.99);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 91.57,135.21) -- ( 91.57,249.68);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 91.57,299.79) -- ( 91.57,283.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.37,135.21) -- ( 96.76,135.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.37,299.79) -- ( 96.76,299.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.18,249.68) --
	(101.96,249.68) --
	(101.96,283.05) --
	( 81.18,283.05) --
	( 81.18,249.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,310.32) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,305.59) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,333.72) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,309.85) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,303.48) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,308.16) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,301.46) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,309.43) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,313.31) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,313.68) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,328.71) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,322.14) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (159.11,282.34) -- (179.89,282.34);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (169.50,206.32) -- (169.50,259.43);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (169.50,307.99) -- (169.50,291.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.31,206.32) -- (174.70,206.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.31,307.99) -- (174.70,307.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.11,259.43) --
	(179.89,259.43) --
	(179.89,291.80) --
	(159.11,291.80) --
	(159.11,259.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,316.59) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,316.53) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (237.05,248.35) -- (257.83,248.35);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (247.44,145.15) -- (247.44,230.90);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (247.44,273.76) -- (247.44,258.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.24,145.15) -- (252.64,145.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.24,273.76) -- (252.64,273.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.05,230.90) --
	(257.83,230.90) --
	(257.83,258.23) --
	(237.05,258.23) --
	(237.05,230.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,285.84) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,276.97) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,277.81) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,275.10) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (340.96,244.14) -- (361.75,244.14);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (351.35,184.73) -- (351.35,235.68);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (351.35,264.95) -- (351.35,252.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (346.16,184.73) -- (356.55,184.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (346.16,264.95) -- (356.55,264.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.96,235.68) --
	(361.75,235.68) --
	(361.75,252.92) --
	(340.96,252.92) --
	(340.96,235.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,277.50) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,272.29) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,273.68) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,269.61) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,274.89) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,271.79) circle (  4.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08, 70.00) -- ( 52.08,279.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08, 70.00) -- ( 43.68, 70.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,174.65) -- ( 43.68,174.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,279.31) -- ( 43.68,279.31);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32, 70.00) {1e-07};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32,174.65) {1e-05};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at (13,210) {$\varepsilon_{\underline{S}_u}$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32,279.31) {1e-03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at (234.45, 10.08) {inputs};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 20.16,198.31) { };
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08, 52.08) --
	(416.82, 52.08) --
	(416.82,344.55) --
	( 52.08,344.55) --
	( 52.08, 52.08);
\end{scope}
\begin{scope}
\path[clip] ( 52.08, 52.08) rectangle (416.82,344.55);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 1.2pt,line join=round] (107.15,280.40) -- (127.94,280.40);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (117.55,198.54) -- (117.55,263.82);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (117.55,313.05) -- (117.55,296.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.35,198.54) -- (122.74,198.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.35,313.05) -- (122.74,313.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.15,263.82) --
	(127.94,263.82) --
	(127.94,296.88) --
	(107.15,296.88) --
	(107.15,263.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,322.21) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,318.08) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,324.66) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,328.89) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,317.19) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (185.09,247.91) -- (205.87,247.91);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (195.48,180.39) -- (195.48,231.77);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (195.48,277.51) -- (195.48,260.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.29,180.39) -- (200.68,180.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.29,277.51) -- (200.68,277.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.09,231.77) --
	(205.87,231.77) --
	(205.87,260.90) --
	(185.09,260.90) --
	(185.09,231.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.48,278.43) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.48,281.93) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.48,282.68) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (263.03,145.96) -- (283.81,145.96);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (273.42, 62.91) -- (273.42,129.15);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (273.42,172.18) -- (273.42,157.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.22, 62.91) -- (278.61, 62.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.22,172.18) -- (278.61,172.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.03,129.15) --
	(283.81,129.15) --
	(283.81,157.95) --
	(263.03,157.95) --
	(263.03,129.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,187.97) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,177.67) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,175.37) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,177.46) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,177.40) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,181.20) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (366.94,147.77) -- (387.72,147.77);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (377.33, 64.25) -- (377.33,127.04);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (377.33,177.11) -- (377.33,159.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.14, 64.25) -- (382.53, 64.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.14,177.11) -- (382.53,177.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.94,127.04) --
	(387.72,127.04) --
	(387.72,159.91) --
	(366.94,159.91) --
	(366.94,127.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,179.32) circle (  4.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.56, 52.08) -- (364.34, 52.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.56, 52.08) -- (104.56, 43.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.49, 52.08) -- (182.49, 43.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.43, 52.08) -- (260.43, 43.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.34, 52.08) -- (364.34, 43.68);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 97.19, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (106.05, 30.24) {3};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (175.12, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (183.99, 30.24) {4};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (253.06, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (261.92, 30.24) {5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (356.97, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (365.84, 30.24) {6};
\end{scope}
\begin{scope}
\path[clip] ( 52.08, 52.08) rectangle (416.82,344.55);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,102.48) rectangle (170.41, 52.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.53, 82.53) rectangle ( 67.83, 88.83);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.53, 65.73) rectangle ( 67.83, 72.03);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 77.28, 79.89) {Variant A.a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 77.28, 63.09) {Variant A.b};
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,315.89) -- (416.82,315.89);
\end{scope}
\end{tikzpicture}

\end{document}