### 2016-09-28 article/Images/boxplots_bis.tex

% Created by tikzDevice version 0.10.1 on 2016-09-28 10:42:51
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,361.35);
\begin{scope}
\path[clip] ( 52.08, 52.08) rectangle (416.82,344.55);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.2pt,line join=round] ( 81.18,306.99) -- (101.96,306.99);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 91.57,276.20) -- ( 91.57,299.28);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 91.57,323.16) -- ( 91.57,311.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.37,276.20) -- ( 96.76,276.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.37,323.16) -- ( 96.76,323.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.18,299.28) --
	(101.96,299.28) --
	(101.96,311.51) --
	( 81.18,311.51) --
	( 81.18,299.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.57,211.08) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (159.11,244.65) -- (179.89,244.65);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (169.50,116.49) -- (169.50,220.17);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (169.50,281.05) -- (169.50,265.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.31,116.49) -- (174.70,116.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.31,281.05) -- (174.70,281.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.11,220.17) --
	(179.89,220.17) --
	(179.89,265.88) --
	(159.11,265.88) --
	(159.11,220.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,320.13) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,321.58) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,321.64) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,327.86) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,320.52) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,333.72) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,322.23) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,320.78) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,333.62) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,332.53) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,324.25) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.50,329.22) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (237.05,262.88) -- (257.83,262.88);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (247.44,115.74) -- (247.44,254.43);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (247.44,302.17) -- (247.44,281.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.24,115.74) -- (252.64,115.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.24,302.17) -- (252.64,302.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.05,254.43) --
	(257.83,254.43) --
	(257.83,281.63) --
	(237.05,281.63) --
	(237.05,254.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,310.30) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,307.71) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,310.68) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,309.44) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,326.37) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,314.83) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,306.55) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.44,319.47) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (340.96,264.78) -- (361.75,264.78);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (351.35, 66.01) -- (351.35,239.90);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (351.35,301.24) -- (351.35,276.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (346.16, 66.01) -- (356.55, 66.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (346.16,301.24) -- (356.55,301.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.96,239.90) --
	(361.75,239.90) --
	(361.75,276.83) --
	(340.96,276.83) --
	(340.96,239.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,305.41) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,308.76) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,306.75) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,314.02) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.35,316.83) circle (  4.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08, 66.54) -- ( 52.08,326.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08, 66.54) -- ( 43.68, 66.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,127.15) -- ( 43.68,127.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,153.25) -- ( 43.68,153.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,213.85) -- ( 43.68,213.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,239.95) -- ( 43.68,239.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,300.56) -- ( 43.68,300.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,326.66) -- ( 43.68,326.66);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32, 66.54) {1e-05};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32,127.15) {5e-05};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at (13,175) {$\varepsilon_{\underline{S}_u}$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32,213.85) {5e-04};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 40.32,300.56) {5e-03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at (234.45, 10.08) {inputs};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 20.16,198.31) { };
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08, 52.08) --
	(416.82, 52.08) --
	(416.82,344.55) --
	( 52.08,344.55) --
	( 52.08, 52.08);
\end{scope}
\begin{scope}
\path[clip] ( 52.08, 52.08) rectangle (416.82,344.55);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 1.2pt,line join=round] (107.15,261.10) -- (127.94,261.10);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (117.55, 62.91) -- (117.55,236.24);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (117.55,302.05) -- (117.55,276.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.35, 62.91) -- (122.74, 62.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.35,302.05) -- (122.74,302.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (107.15,236.24) --
	(127.94,236.24) --
	(127.94,276.57) --
	(107.15,276.57) --
	(107.15,236.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,303.09) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,307.34) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.55,314.20) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (185.09,248.49) -- (205.87,248.49);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (195.48, 75.27) -- (195.48,227.29);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (195.48,297.89) -- (195.48,271.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.29, 75.27) -- (200.68, 75.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.29,297.89) -- (200.68,297.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.09,227.29) --
	(205.87,227.29) --
	(205.87,271.55) --
	(185.09,271.55) --
	(185.09,227.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.48,301.86) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (263.03,195.72) -- (283.81,195.72);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (273.42, 76.18) -- (273.42,169.85);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (273.42,230.87) -- (273.42,207.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.22, 76.18) -- (278.61, 76.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.22,230.87) -- (278.61,230.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.03,169.85) --
	(283.81,169.85) --
	(283.81,207.60) --
	(263.03,207.60) --
	(263.03,169.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,243.30) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,245.14) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,251.06) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,235.07) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,243.94) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.42,232.91) circle (  4.41);

\path[draw=drawColor,line width= 1.2pt,line join=round] (366.94,193.08) -- (387.72,193.08);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (377.33, 73.03) -- (377.33,172.11);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (377.33,239.43) -- (377.33,216.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.14, 73.03) -- (382.53, 73.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.14,239.43) -- (382.53,239.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.94,172.11) --
	(387.72,172.11) --
	(387.72,216.80) --
	(366.94,216.80) --
	(366.94,172.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,253.83) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,253.84) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,262.95) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,270.28) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,247.69) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,254.87) circle (  4.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.33,251.83) circle (  4.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.56, 52.08) -- (364.34, 52.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.56, 52.08) -- (104.56, 43.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.49, 52.08) -- (182.49, 43.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.43, 52.08) -- (260.43, 43.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.34, 52.08) -- (364.34, 43.68);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 97.19, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (106.05, 30.24) {3};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (175.12, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (183.99, 30.24) {4};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (253.06, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (261.92, 30.24) {5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at (356.97, 32.77) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.18] at (365.84, 30.24) {6};
\end{scope}
\begin{scope}
\path[clip] ( 52.08, 52.08) rectangle (416.82,344.55);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,102.48) rectangle (170.41, 52.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.53, 82.53) rectangle ( 67.83, 88.83);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.53, 65.73) rectangle ( 67.83, 72.03);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 77.28, 79.89) {Variant A.a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.68] at ( 77.28, 63.09) {Variant A.b};
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.08,300.56) -- (416.82,300.56);
\end{scope}
\end{tikzpicture}

\end{document}